---
title: ""
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

> This repository stores the code used to perform the GLM analyses presented in the manuscript _Evaluating the long-term net reproductive benefits and costs of supplementation in upper Yakima River Chinook salmon_ by I.J. Koch, B.A. Staton, H.M. Nuetzel, T.R. Seamons, A.P. Matala, K.I. Warheit, M.V. Johnston, C.R. Strom, S.R. Narum, and W.J. Bosch.

[![ArticleDOI](https://img.shields.io/badge/Article-PLACEHOLDER%20IF%20ACCEPTED-blue?logo=doi&logoColor=f5f5f5)]()  
[![GitHub Repo Archive DOI](https://img.shields.io/badge/GitHub%20Repo%20Archive-PLACEHOLDER%20WHEN%20MINTED-blue?logo=github)]()

## Data Ownership

_The data file needed to run these analyses has not yet been approved for public sharing.
Please contact the manuscript authors to receive a copy of the data set._

## Repo Organization

This repository contains an assortment of GLM analyses intended to address research questions about the relative reproductive success of spawning Chinook salmon in the upper Yakima River.

```{r subdir-table}
tab = matrix(c(
  "`01-1gen-RRS`",      "Compares progeny produced by HOR and NOR spawners reproducing naturally.",
  "`02a-2gen-RRS`",     "Compares grand-progeny produced by HOR and NOR spawners reproducing naturally.",
  "`02b-2gen-RRS-wF1`", "Compares grand-progeny produced by HOR and NOR spawners reproducing naturally, while considering the number of progeny produced as a predictor.",
  "`03-cross-type-RRS`",  "Compares progeny produced by spawning pairs in which the parents are of differing origin types.",
  "`04-acc-site-RRS`",    "Compares progeny produced by HOR spawners reproducing naturally based on the site they were acclimated at as smolts.",
  "`05-ancestry-RRS`",    "Compares progeny produced by NOR spawners reproducing naturally based on the origin types of their parents.",
  "`06-1gen-demo-boost`", "Compares progeny produced by NOR spawners based on where they reproduced -- either naturally or in the hatchery as broodstock.",
  "`07-2gen-demo-boost`", "Compares grand-progeny produced by NOR spawners based on where they reproduced -- either naturally or in the hatchery as broodstock.",
  "`08-attr-v-origin`",   "Compares length and return date by origin."
), ncol = 2, byrow = TRUE)

knitr::kable(tab, "markdown", col.names = c("Subdirectory", "Description"))

```

Additional scripts include:

* `common-functions.R`: Stores custom functions used to streamline performing replicate tasks across multiple analyses.
* `run-one.R`: Script to execute one of the analyses shown in the table above.
* `run-all.R`: Script to execute all of the analyses shown in the table above, and build manuscript figures, tables, and supplemental material.
* `ms-content/ms-figs.R`: Script to build manuscript figures that rely on the output of GLM analyses.
* `ms-content/ms-tables.Rmd`: Rmarkdown to build manuscript tables.
* `ms-content/supplement`: Contains multiple Rmarkdown source code files for building the manuscript supplemental material.

## Reproducibility

The output files are too large to track in a git repo, so this code will need to be executed locally to reproduce the output presented in the manuscript and supplement. 

All analyses that use random number generators (e.g., parametric bootstrapping) include a `set.seed()` statement, so our results should be exactly reproducible if all defaults are used.

Clone the repository, acquire the data (see _Data Ownership_ above), and navigate to the repository location on your computer.

**All Analyses At Once**

The exact output presented in the manuscript can be most easily reproduced via command line, e.g.:

```bash
Rscript run-all.R
```

Or from within R:

```R
source("run-all.R")
```

This will run each of the GLM analyses in the table above sequentially, then build the manuscript figures and tables and the supplement. 
When complete, the output will be found the `ms-content/output` subdirectory, and in the `/output` subdirectories of each analysis-specific directory.

>_The default is to use 8 CPU cores to run 1,000 bootstrap iterations per GLM analysis. 
>This takes approximately 9 hours on a current modern laptop computer (RAM: 64GB; Processor: 20 threads, 5.40 GHz Max Turbo, 24 MB Cache).
>Similar results to those we present can be obtained with fewer bootstrap iterations -- change the value of the `nboot` variable in the `run-all.R` script to, e.g., 100 to obtain results faster._

**One-by-One**

Any of the specific GLM analyses can be executed with all defaults, e.g.,

```bash
Rscript run-one.R 05-ancestry-RRS
```

Or execute the script with the `-h` flag to see how to change the defaults:

```bash
Rscript run-one.R -h
```

**Interactively**

Of course, you can open the `yakima-RRS.Rproj` file in [RStudio Desktop](https://posit.co/download/rstudio-desktop/) and run the scripts in each of the subdirectories interactively to gain insights about how the code works.

## Dependencies

The table below shows all R packages (and their versions at time of publication) that are specifically called (typically with `pkg::fn()`) or loaded (using `library(pkg)`).

```{r pkg-table}

# detect all packages called specifically
used_pkgs = renv::dependencies(quiet = TRUE)$Package |> 
  unique() |> 
  sort()

# exclude any packages you don't wish to include in the table here
# e.g., anything used solely in building this Rmd
used_pkgs = used_pkgs[!used_pkgs %in% c("renv", "sessioninfo")]

# get the version of each package
versions = sapply(used_pkgs, function(pkg) paste(packageVersion(pkg), collapse = "."))

# get the description (i.e., title field of DESCRIPTION file) of each package
desc = sapply(used_pkgs, function(pkg) paste(packageDescription(pkg)$Title, collapse = ".")) |> 
  stringr::str_replace_all("\\n", " ")

# build a matrix
tab = cbind(pkg = used_pkgs, version = versions, desc = desc)
rownames(tab) = NULL

# build package name as a markdown link, with a URL placeholder
tab[,"pkg"] = paste0("[`", tab[,"pkg"], "`](URL)")

# function to get the URL that will replace the placeholder
get_url = function(pkg) {
  paste0("https://CRAN.R-project.org/package=", pkg)
}

# get the URLs for each package and perform the replacement
tab[,"pkg"] = stringr::str_replace(tab[,"pkg"], "URL", sapply(used_pkgs, get_url))

# return the kable
kableExtra::kbl(tab, "markdown", col.names = c("Package", "Version", "Description"), align = "lrl")
```

Running the code below will display the packages not already installed on your machine:

```{r, eval = TRUE, echo = FALSE}
check_code = paste0("```R\npkgs <- c(", knitr::combine_words(used_pkgs, before = '"', and = "", sep = ', '), ")\npkgs[!pkgs %in% rownames(installed.packages())]\n```")
```

`r check_code`

## Session Info

<details>
<summary><b>Click to Expand/Hide Session Info</b></summary>

```{r message = FALSE, warning = FALSE}
junk = suppressPackageStartupMessages({
  sapply(used_pkgs, library, character.only = TRUE)
})
```

```{r}
sessioninfo::session_info()
```

</details>
