---
title: ""
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

> This repository stores the code used to perform the GLM analyses presented in the manuscript _Evaluating the long-term net reproductive benefits and costs of supplementation in upper Yakima River Chinook salmon_ by I.J. Koch, B.A. Staton, H.M. Nuetzel, T.R. Seamons, A.P. Matala, K.I. Warheit, M.V. Johnston, C.R. Strom, S.R. Narum, and W.J. Bosch; now accepted for publication in the _Canadian Journal of Fisheries and Aquatic Sciences_.

[![ArticleDOI](https://img.shields.io/badge/Article-10.1139/cjfas--2024--0289-blue?logo=doi&logoColor=f5f5f5)](https://doi.org/10.1139/cjfas-2024-0289)  
[![GitHub Repo Archive DOI](https://img.shields.io/badge/GitHub%20Repo%20Archive-10.5281/zenodo.13738201-blue?logo=github)](https://doi.org/10.5281/zenodo.13738201)

## Repo Organization

This repository contains an assortment of GLM analyses intended to address research questions about the relative reproductive success of spawning Chinook salmon in the upper Yakima River.

```{r subdir-table}
tab = matrix(c(
  "`01-1gen-RRS`",      "Compares progeny produced by HOR and NOR spawners reproducing naturally.",
  "`02a-2gen-RRS`",     "Compares grand-progeny produced by HOR and NOR spawners reproducing naturally.",
  "`02b-2gen-RRS-wF1`", "Compares grand-progeny produced by HOR and NOR spawners reproducing naturally, while considering the number of progeny produced as a predictor.",
  "`03-cross-type-RRS`",  "Compares progeny produced by spawning pairs in which the parents are of differing origin types.",
  "`04-acc-site-RRS`",    "Compares progeny produced by HOR spawners reproducing naturally based on the site they were acclimated at as smolts.",
  "`05-ancestry-RRS`",    "Compares progeny produced by NOR spawners reproducing naturally based on the origin types of their parents.",
  "`06-1gen-percapita-prod`", "Compares progeny produced by NOR spawners based on where they reproduced -- either naturally or in the hatchery as broodstock.",
  "`07-2gen-percapita-prod`", "Compares grand-progeny produced by NOR spawners based on where they reproduced -- either naturally or in the hatchery as broodstock.",
  "`08-attr-v-origin`",   "Compares length and return date by origin."
), ncol = 2, byrow = TRUE)

knitr::kable(tab, "markdown", col.names = c("Subdirectory", "Description"))

```

Additional scripts include:

* `common-functions.R`: Stores custom functions used to streamline performing replicate tasks across multiple analyses.
* `run-one.R`: Script to execute one of the analyses shown in the table above.
* `run-all.R`: Script to execute all of the analyses shown in the table above, and build manuscript figures, tables, and supplemental material.
* `ms-content/ms-figs.R`: Script to build manuscript figures that rely on the output of GLM analyses.
* `ms-content/ms-tables.Rmd`: Rmarkdown to build manuscript tables.
* `ms-content/supplement`: Contains multiple Rmarkdown source code files for building the manuscript supplemental material.

## Data File

The data file containing the records for all GLM analyses is `YakimaParentageInputFile2024.xlsx`.
It has two tabs, which contain data for the different analyses, click below to view a description of each variable.

<details>
  <summary><b>Tab: Yakima Assign Per Cross</b></summary>

Each row is a detected spawning pair.

```{r}
tab = matrix(c(
  "`Pa_Ma`",                    "Individual IDs of the male and female spawners involved in the pair (separated by a colon)",
  "`No.Offspring`",             "Number of adult offspring that were assigned to the pair",
  "`Parental_CrossType`",       "Origin of the male and female parent; expressed as `maleXfemale` (`WN` = naturally spawned, `SH` = hatchery spawned)",
  "`Parental_Spawn_Location`",  "Setting in which the spawning occurred for this pair, either natural or hatchery",
  "`Parental_SampleYear`",      "Year in which spawning occurred",
  "`Pa_Ma_life_stage`",         "Life stages (either jack or adult) of the male and female parent; expressed as `male_female`; _all female jacks recoded as adults for analysis_",
  "`Pa_Disposition`",           "Spawning disposition of the male parent (natural or broodstock)",
  "`Pa_Origin`",                "Origin of the male parent (`WN` = naturally spawned, `SH` = hatchery spawned)",
  "`Pa_Date`",                  "Return date of the male parent",
  "`Pa_Location`",              "Spawning location of the male parent (natural or hatchery)",
  "`Pa_ForkLength`",            "Fork length (mm) of the male parent",
  "`Ma_Disposition`",           "Spawning disposition of the female parent (natural or broodstock)",
  "`Ma_Origin`",                "Origin of the female parent (`WN` = naturally spawned, `SH` = hatchery spawned)",
  "`Ma_Date`",                  "Return date of the female parent",
  "`Ma_Location`",              "Spawning location of the female parent (natural or hatchery)",
  "`Ma_ForkLength`",            "Fork length (mm) of the female parent"
  
), ncol = 2, byrow = TRUE)

knitr::kable(tab, "markdown", col.names = c("Column", "Description"))
```

</details>

<details>
  <summary><b>Tab: YakimaAssignmentsGrand</b></summary>

Each row is an individual spawner.

```{r}
tab = matrix(c(
  "`id`",                      "Individual spawner ID",
  "`sex`",                     "Sex of the spawner",
  "`origin`",                  "Origin of the spawner (i.e., `WN` = naturally spawned, `SH` = hatchery spawned)",
  "`length`",                  "Fork length of the spawner (mm)",
  "`date`",                    "Return date of the spawner",
  "`day`",                     "Return date, converted to day of the year",
  "`disposition`",             "Spawning disposition (natural or broodstock)",
  "`life_stage`",              "Life stage of the spawner (jack or adult; _all female jacks recoded as adult for analysis_)",
  "`total_adult_offs`",        "Number of adult offspring that were assigned to the spawner",
  "`Total_Adult_Grand_Offs`",  "Number of adult grand-offspring (i.e., second generation, offspring's offspring) assigned to the spawner",
  "`Parentage Broodyear`",     "Year in which the spawner returned to spawn",
  "`Recoded Acc Site`",        "For hatchery-origin fish, the acclimation site to which they were released",
  "`Pa_Ma`",                   "Where available, the IDs of the parents of the spawner, expressed as `male:female`"

), ncol = 2, byrow = TRUE)

knitr::kable(tab, "markdown", col.names = c("Column", "Description"))
```

</details>

## Reproducibility

The output files are too large to track in a git repo, so this code will need to be executed locally to reproduce the output presented in the manuscript and supplement. 

All analyses that use random number generators (e.g., parametric bootstrapping) include a `set.seed()` statement, so our results should be exactly reproducible if all defaults are used.

Clone the repository and navigate to the repository location on your computer.

**All Analyses At Once**

The exact output presented in the manuscript can be most easily reproduced via command line, e.g.:

```bash
Rscript run-all.R
```

Or from within R:

```R
source("run-all.R")
```

This will run each of the GLM analyses in the table above sequentially, then build the manuscript figures and tables and the supplement. 
When complete, the output will be found the `ms-content/output` subdirectory, and in the `/output` subdirectories of each analysis-specific directory.

>_The default is to use 10 CPU cores to run 1,000 bootstrap iterations per GLM analysis. 
>This takes approximately 6 hours on a modern laptop computer (RAM: 64GB; Processor: 20 threads, 5.40 GHz Max Turbo, 24 MB Cache).
>Similar results to those we present can be obtained with fewer bootstrap iterations -- change the value of the `nboot` variable in the `run-all.R` script to, e.g., 100 to obtain results faster._

**One-by-One**

Any of the specific GLM analyses can be executed with all defaults, e.g.,

```bash
Rscript run-one.R 05-ancestry-RRS
```

Or execute the script with the `-h` flag to see how to change the defaults:

```bash
Rscript run-one.R -h
```

**Interactively**

Of course, you can open the `yakima-RRS.Rproj` file in [RStudio Desktop](https://posit.co/download/rstudio-desktop/) and run the scripts in each of the subdirectories interactively to gain insights about how the code works.

## Dependencies

The table below shows all R packages (and their versions at time of publication) that are specifically called (typically with `pkg::fn()`) or loaded (using `library(pkg)`).

```{r pkg-table}

# detect all packages called specifically
used_pkgs = renv::dependencies(quiet = TRUE)$Package |> 
  unique() |> 
  sort()

# exclude any packages you don't wish to include in the table here
# e.g., anything used solely in building this Rmd
used_pkgs = used_pkgs[!used_pkgs %in% c("renv", "sessioninfo")]

# get the version of each package
versions = sapply(used_pkgs, function(pkg) paste(packageVersion(pkg), collapse = "."))

# get the description (i.e., title field of DESCRIPTION file) of each package
desc = sapply(used_pkgs, function(pkg) paste(packageDescription(pkg)$Title, collapse = ".")) |> 
  stringr::str_replace_all("\\n", " ")

# build a matrix
tab = cbind(pkg = used_pkgs, version = versions, desc = desc)
rownames(tab) = NULL

# build package name as a markdown link, with a URL placeholder
tab[,"pkg"] = paste0("[`", tab[,"pkg"], "`](URL)")

# function to get the URL that will replace the placeholder
get_url = function(pkg) {
  paste0("https://CRAN.R-project.org/package=", pkg)
}

# get the URLs for each package and perform the replacement
tab[,"pkg"] = stringr::str_replace(tab[,"pkg"], "URL", sapply(used_pkgs, get_url))

# return the kable
kableExtra::kbl(tab, "markdown", col.names = c("Package", "Version", "Description"), align = "lrl")
```

Running the code below will display the packages not already installed on your machine:

```{r, eval = TRUE, echo = FALSE}
check_code = paste0("```R\npkgs <- c(", knitr::combine_words(used_pkgs, before = '"', and = "", sep = ', '), ")\npkgs[!pkgs %in% rownames(installed.packages())]\n```")
```

`r check_code`

## Session Info

<details>
<summary><b>Click to Expand/Hide Session Info</b></summary>

```{r message = FALSE, warning = FALSE}
junk = suppressPackageStartupMessages({
  sapply(used_pkgs, library, character.only = TRUE)
})
```

```{r}
sessioninfo::session_info()
```

</details>
