#### Model Coefficients {.tabset .tabset-pills}

:::{.helptext}
<details>
<summary>`r shiny::icon("info-circle")` **Description**</summary>
_These tables display the estimated coefficients of the GLM submodels._

_For the **conditional submodel**, the intercept represents $\log(\mu)$ at zero/baseline values of all predictors, where $\mu$ is the mean parameter of a zero-truncated negative binomial distribution; other coefficients represent the additive effect of changes in the covariates on $\log(\mu)$ (i.e., multiplicative effects on $\mu$)._

_For the **zero submodel**, the intercept represents $\log(\pi/(1-\pi))$ (logit transformation, a.k.a. log-odds) at zero/baseline values of all predictors, where $\pi$ represents the probability the response is zero; other coefficients represent the additive effect of changes in the covariates on $\log(\pi/(1-\pi))$ (i.e., multiplicative effects on odds of a zero)._
</details>
:::

##### Conditional Submodel

```{r}
coef_kable(best_model, type = "cond")
```

`r if (has_zi) "##### Zero Submodel"`

```{r, eval = has_zi}
coef_kable(best_model, type = "zi")
```
